---
title: "Appendix: Comparison of decomposition methods"
author: Rob J Hyndman and Alexander Dokumentov
bibliography: strrefs.bib
biblio-style: authoryear-comp
fontfamily: mathpazo
fontsize: 11pt
output:
  MonashEBSTemplates::memo:
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    toc: no
    keep_tex: yes
    number_sections: yes
header-includes:
  - \graphicspath{{figs/}}
  - \def\addlinespace{}
  - \usepackage{amsmath}
  - \usepackage{carlito}
  - \setlength{\arraycolsep}{3pt}
  - \setcounter{MaxMatrixCols}{20}
  - \newcommand{\argmin}{\operatorname{arg\,min}}
  - \newcommand{\sn}{\kappa}
  - \renewcommand{\vec}{\operatorname{vec}}
  - \newcommand{\diag}{\operatorname{diag}}
  - \renewcommand{\mod}{\operatorname{mod}}
  - \usepackage[section]{placeins}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning=FALSE,
  message=FALSE,
  cache=TRUE
)
library(tidyverse)
library(stR)
library(fabletools)
library(distributional)
library(lubridate)
library(forecast)
library(imputeTS)
library(tsibble)

# Create predictor dist object
predictor_dist <- function(x, level) {
    dist_percentile(
      with(x, mapply(c, data, split(lower, row(lower)), split(upper, row(upper)), SIMPLIFY = FALSE)),
      list(c(50, 50 + rep(c(-1, 1), each = length(level))*rep(level, 2)/2))
    )
}
```

We compare STR with three other decomposition methods on monthly data of supermarket and grocery store log revenue in New South Wales, Australia. The three comparison methods are STL [@cleveland1990stl], TBATS [@delivera2011forecasting], and X-13-ARIMA-SEATS [@findley2005some;@Dagum2016]

```{r realdata, fig.height=4.5, fig.cap="Supermarket and grocery store log revenue in New South Wales"}
supermarket_nsw <- tsibbledata::aus_retail %>%
  filter(
    State == "New South Wales",
    Industry == "Supermarket and grocery stores",
    year(Month) >= 2000,
    year(Month) <= 2009
  ) %>%
  pull(Turnover) %>% 
  log() %>%
  ts(start=2000, frequency=12)
# STR decomposition
d <- supermarket_nsw %>%
  AutoSTR(confidence=0.95, gapCV=12, reltol=0.00001)

# STL decomposition
d_stl = forecast::mstl(supermarket_nsw, iterate=5)

# TBATS decomposition
d_tbats = tbats(supermarket_nsw, use.box.cox=FALSE, use.trend=TRUE)

# X-13-ARIMA-SEATS decomposition
d_x13 <- seasonal::seas(supermarket_nsw, transform.function = "none")

# Make data frame with components
decomp <- bind_rows(
    tibble(
      method = "STR",
      trend = d$output$predictors[[1]]$data,
      seasonal = d$output$predictors[[2]]$data,
      remainder = d$output$random$data,
      month = yearmonth(seq(from=as.Date("2000-01-01"),by="1 month", length=120))
    ),
    as_tibble(d_stl) %>%
      transmute(
        method = "STL",
        trend=Trend, 
        seasonal=Seasonal12, 
        remainder=Remainder,
        month = yearmonth(seq(from=as.Date("2000-01-01"),by="1 month", length=120))
      ),
    tbats.components(d_tbats) %>%
      as_tibble() %>%
      transmute(
        method = "TBATS",
        trend = level,
        seasonal = season,
        remainder = observed - level - season,
      month = yearmonth(seq(from=as.Date("2000-01-01"),by="1 month", length=120))
      ),
    tibble(
      method = "X-13-ARIMA-SEATS",
      trend = seasonal::trend(d_x13),
      #seasonal = seasonal::series(d_x13, "seats.seasonal"),
      remainder = seasonal::irregular(d_x13),
      seasonal = supermarket_nsw - trend - remainder,
      month = yearmonth(seq(from=as.Date("2000-01-01"),by="1 month", length=120))
    )
  ) %>%
  mutate(observed = trend + seasonal + remainder)
# Check
if(
  sum(abs(filter(decomp, method=="STR")$observed - filter(decomp, method=="STL")$observed)) +
  sum(abs(filter(decomp, method=="STR")$observed - filter(decomp, method=="TBATS")$observed)) +
  sum(abs(filter(decomp, method=="STR")$observed - filter(decomp, method=="X-13-ARIMA-SEATS")$observed)) > 1e-10)
  stop("Components not adding up")
```

The components are shown below.

```{r plot, fig.asp=1.1}
decomp %>%
  select(-observed) %>%
  bind_rows(tsibble(
    method="Data",
    data = supermarket_nsw,
    month = yearmonth(seq(from=as.Date("2000-01-01"),by="1 month", length=120)),
    index=month, key=method
  )) %>%
  select(month, method, everything()) %>%
  pivot_longer(data:remainder, 
               names_to = "variable", values_to="value") %>%
  mutate(
    variable = factor(variable, levels=c("data","trend","seasonal","remainder")),
    method = factor(method, levels=c("Data","STR","STL","TBATS","X-13-ARIMA-SEATS"))
  ) %>%
  ggplot(aes(x=month, y=value, group=method, col=method)) +
  geom_line() +
  facet_grid(variable ~ ., scales = "free_y") +
  scale_color_manual(values = c(Data = "black",STL="#D55E00",STR="#0072B2",TBATS="#009E73",
                                `X-13-ARIMA-SEATS` = "#CC79A7")) +
  guides(colour = guide_legend(title = "  ")) +
  labs(y="", x="Month")
```

The results are very similar, with the main differences as follows.

  * The trend component obtained via TBATS is less smooth than the other methods. This is not surprising, as TBATS uses only data prior to each time period to estimate the components, whereas the other methods use all the data to estimate components.
  * X-13-ARIMA-SEATS allows the seasonal component to change more rapidly than the other methods, resulting in a smaller remainder component. Essentially some of the variation in the remainder component has been incorporated into the seasonal component (and to a lesser extent into the trend component). In our judgement, this is over-fitting the data.
